6.867 Problem Set 1

### 2015-12-10 Project/Latex/Paper.tex
[project] first pass at regression discussion.

@@

>>>>

Once we have constructed our full set of features, we aim to predict log(article pageviews), $\mathbf{y}$, using our full set of variables, which we include in our design matrix, $\mathbf{X}$. We estimate the feature weights using the closed form solution for both OLS and ridge regression,

\begin{center}
\begin{equation}
\beta = (\mathbf{X}^T\mathbf{X} + \lambda \mathbf{I})^{-1}\mathbf{X}^T\mathbf{y}
\end{equation}
\end{center}

\noindent where $\mathbf{I}$ is the $k \times k$ identity matrix, and $\lambda$ is our regularization parameter. Setting $\lambda = 0$ corresponds to OLS, whereas non-zero values of $\lambda$ correspond to ridge regression. The motivation for performing ridge regression as opposed to OLS is to not overfit on our data.

In order to choose an appropriate value of $\lambda$, we split our data into a training and validation set. 90\% of the data is allocated to the training set, and 10\% of the data is allocated to the test set. We cross-validate the estimated $\beta$ values on our validation set for each $\lambda$, and choose the value of $\lambda$ that produces the lowest MSE on the validation data.

The table below displays the 20 feature weights with the largest magnitudes given our model. A bar chart showing the magnitude of all of the weights (excluding the intercept weight) can be found in Figure \ref{fig:lin_reg_weights}.

@@

Figure \ref{fig:mse_plot} shows the training and holdout MSE for various values of Lambda. There are a few things in this plot worth discussing. First, note that the holdout MSE is consistently lower than the training data MSE. Given the (relatively) small size of our dataset (6,687 observations), this is likely due to the sampling we used to separate our data into training and validation data. However, this shouldn't effect the validity of our cross validation.

Another thing worth noting is that even on the training dataset, the MSE goes down once we choose a non-zero value of $\lambda$. At first, this was surprising to our group, as conceptually OLS is often thought of as the linear regression method that minimizes MSE. However, it is important to note that OLS only holds this distinction amongst unbiased estimators. Hoerl and Kennard (1970) \cite{hoerl1970ridge} prove the existence theorem for ridge regression, which proves that there exists a value of $\lambda$ for which $\beta_{ridge}$ produces a lower MSE than $\beta_{OLS}$.

Another way of framing this finding is through bias-variance tradeoff. Recall that the MSE can be written as a function of the bias and variance:

\begin{equation}
MSE = (\textrm{Bias})^2 + \textrm{Var}.
\end{equation}

\noindent For some values of $\lambda$, ridge regression is able to lower the MSE by decreasing variance, but adding a non-zero bias term. We believe this is what is happening in our data when cycling through different $\lambda$ values.
